---
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: false
    css: "brand_styles.css"
    includes:
      before_body: "before_body.html"
    self_contained: false
---

```{r intro-block, results='asis', echo=FALSE}

htmltools::div(
  class = "intro-wrapper",
  
  htmltools::div(
    style = "
      position:relative;
      margin-top:18px;
      display:grid;
      grid-template-columns: 2.2fr 1fr;
      gap:20px;
      align-items:stretch;
    ",
    
  # ================= LEFT PANEL =================
  htmltools::div(
style = "
  background: rgba(25, 95, 170, 0.55);
  backdrop-filter: blur(14px);
  -webkit-backdrop-filter: blur(14px);
  border: 1px solid rgba(255,255,255,0.16);
  border-radius:18px;
  overflow:hidden;
  box-shadow:0 6px 20px rgba(0,0,0,0.18);
  color:white;
  display:flex;
  flex-direction:column;
",

    # ---------- BANNER IMAGE ----------
    htmltools::tags$img(
      src = "FB TRPP.png",
      style = "
        width:100%;
        height:220px;
        object-fit:cover;
        display:block;
      "
    ),

    # ---------- TEXT CONTENT ----------
    htmltools::div(
      style = "
        padding:22px 24px 24px 24px;
        font-size:15px;
        line-height:1.65;
        color:rgba(255,255,255,0.95);
      ",

      htmltools::tags$p(
        htmltools::a(
          href = "https://www.therockpoolproject.co.uk/the-big-rock-pool-challenge",
          style = "
            color:#D6F2FF;
            font-weight:700;
            text-decoration:none;
          ",
          "The Big Rock Pool Challenge"
        ),
        " is a UK-wide citizen science initiative run by the Rock Pool Project, bringing people together to explore and record life on UK rocky shores. Participants take part in monthly BioBlitz Battle games, recording coastal marine life at designated monitoring hubs using the iNaturalist platform and the BioBlitz Battle app."
      ),

      htmltools::tags$p(
        "Records are verified through community identifications, allowing observations to reach research-grade status. Every record added helps build a growing picture of coastal biodiversity around the UK."
      ),

      htmltools::tags$p(
        "The Big Rock Pool Challenge is run by The Rock Pool Project, in partnership with the Marine Biological Association, and is funded by the ScottishPower Foundation and the National Lottery Heritage Fund."
      )
    )
  ),

  # ================= RIGHT PANEL =================
  htmltools::div(
    style = "
      background:#ffffff;
      color:#222222;
      border-radius:18px;
      padding:20px;
      display:flex;
      flex-direction:column;
      justify-content:center;
      align-items:center;
      gap:24px;
      box-shadow:0 6px 18px rgba(0,0,0,0.10);
    ",

    htmltools::tags$div(
      "Funding & Partners",
      style = "
        font-size:15px;
        font-weight:700;
        letter-spacing:0.3px;
        color:#222222;
      "
    ),

    htmltools::tags$img(
      src = "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/67dc251080bf058c3bffd547_hf-logo.png",
      style = "
        width:100%;
        max-width:220px;
        object-fit:contain;
      "
    ),

    htmltools::tags$img(
      src = "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/690234a2aeabe1446d63c1cc_AAFF%20LOGOTYPE%20SCOTTISHPOWER%20FOUNDATION%20RGB-p-2000.jpg",
      style = "
        width:100%;
        max-width:220px;
        object-fit:contain;
      "
    ),

    htmltools::tags$img(
      src = "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/67b4f7fae1b32ae21a76b697_Untitled%20design%20(11).png",
      style = "
        width:100%;
        max-width:220px;
        object-fit:contain;
      "
    )
  )
)
)

```

::: {style="height:20px;"}
:::

```{=html}
<script>
window.addEventListener("load", function() {

  function moveTOC() {

    var toc = document.getElementById("TOC");
    var target = document.getElementById("toc-here");

    if (toc && target && !target.contains(toc)) {
      target.appendChild(toc);
      toc.style.marginTop = "20px";
    } else {
      setTimeout(moveTOC, 100);
    }
  }

  moveTOC();
});
</script>
```

::: {#toc-here}
:::

```{r setup-chunk, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

library(dplyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(sf)
library(tidyr)
library(vegan)
library(ggrepel)
library(scales)
library(forcats)
library(glue)
library(purrr)
library(rnaturalearth)
library(viridis)
library(htmltools)
library(ggimage)

theme_set(
  theme_minimal() +
    theme(
      legend.position = "right",
      plot.caption = element_text(hjust = 0)
    )
)

safe_id <- function(x) {
  gsub("[^A-Za-z0-9]", "-", x)
}

get_col <- function(df, nm) {
  if (nm %in% names(df)) df[[nm]] else rep(NA_character_, nrow(df))
}

default_avatar <- "https://cdn.prod.website-files.com/6761e748382f7a483dc6c262/68c02913eb9eedda1582018a_iNaturalist%20app.png"

safe_img <- function(x) {
  ifelse(is.na(x) | x == "", default_avatar, x)
}

safe_species_image <- function(photo_url,
                               attribution,
                               license_ok = TRUE) {

  ok <- !is.na(photo_url) &
        photo_url != "" &
        license_ok

  out_photo <- ifelse(ok, photo_url, default_avatar)

  out_attr <- ifelse(ok, attribution, "")

  tibble(
    photo = out_photo,
    attribution = out_attr
  )
}

allowed_licenses <- c(
  "cc-by",
  "cc-by-sa",
  "cc-by-nc",
  "cc-by-nc-sa",
  "cc0"
)


```

```{r data-download, echo=FALSE}

source("../../Ben/iNat API functions.R")

iconic_lookup <- read.csv("../../Data/BRPC_iconic_taxa.csv")

objs <- load("../../Data/rarity_scores2.1.RData")
rarity_scores <- get(objs[1])

rarity_scores <- rarity_scores %>%
  select(-Score) %>%
  rename(Score = Score100)

load("../../Ben/BRPC_dat.RData")
```

```{r hub-setup}

hub_sf <- tibble::tribble(
  ~hub, ~lon, ~lat,
  "Castle Beach", -5.06, 50.15,
  "Lee-on-the-Solent", -1.20, 50.80,
  "Beadnell Haven", -1.63, 55.56,
  "Ovingdean", -0.08, 50.81,
  "Shoeburyness", 0.78, 51.53,
  "Menai Bridge", -4.16, 53.23,
  "Mount Batten", -4.14, 50.36,
  "Kilve Beach", -3.15, 51.20,
  "East Sands", -2.79, 56.34
) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  st_transform(3857)
```

```{r data-pipeline}

# ---------------------------------------
# ALL DATA
# ---------------------------------------

BRPC_dat <- BRPC_dat %>%
  mutate(
    scientific_name = dplyr::coalesce(
      get_col(., "scientific_name"),
      get_col(., "taxon.name"),
      get_col(., "taxon_name"),
      get_col(., "species_guess")
    )
  )
# ---------------------------------------
# PROJECT DATA (up to Q2 end)
# ---------------------------------------

project_end <- as.Date("2026-06-30")

BRPC_dat <- BRPC_dat %>%
  mutate(observed_on = as.Date(observed_on)) %>%
  filter(observed_on <= project_end)

all_sf <- BRPC_dat %>%
  mutate(
    lon = purrr::map_dbl(
      geojson.coordinates,
      ~ if(length(.x) >= 1) .x[[1]] else NA_real_
    ),
    lat = purrr::map_dbl(
      geojson.coordinates,
      ~ if(length(.x) >= 2) .x[[2]] else NA_real_
    )
  ) %>%
  filter(!is.na(lon), !is.na(lat)) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  st_transform(3857)

all_sf$hub <- hub_sf$hub[
  st_nearest_feature(all_sf, hub_sf)
]

BRPC_dat <- all_sf %>%
  st_drop_geometry()

# ---------------------------------------
# Q2 DATA
# ---------------------------------------

Q2_base <- BRPC_dat %>%
  filter(
    observed_on >= "2026-04-01",
    observed_on <= "2026-06-30"
  )

Q2_sf <- Q2_base %>%
  mutate(
    lon = purrr::map_dbl(geojson.coordinates, 1),
    lat = purrr::map_dbl(geojson.coordinates, 2)
  ) %>%
  filter(!is.na(lon), !is.na(lat)) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  st_transform(3857)

Q2_sf$hub <- hub_sf$hub[
  st_nearest_feature(Q2_sf, hub_sf)
]

Q2_base$hub <- Q2_sf$hub
```

```{r helper-functions}

render_summary_card <- function(title, observations, contributors, species) {

  div(
    style = "
      background:#ffffff;
      border-radius:20px;
      padding:26px;
      margin:20px 0 30px 0;
      box-shadow:0 4px 14px rgba(0,0,0,0.08);
      border:1px solid #e8e8e8;
      text-align:center;
    ",

    h2(title),

    div(
      style = "
        display:inline-flex;
        gap:18px;
        flex-wrap:wrap;
        justify-content:center;
        align-items:center;
        background:#f5f5f5;
        padding:14px 22px;
        border-radius:999px;
        margin-top:10px;
      ",

      HTML(glue("
        <span><strong>{comma(observations)}</strong> observations</span>
        <span>·</span>
        <span><strong>{comma(contributors)}</strong> contributors</span>
        <span>·</span>
        <span><strong>{comma(species)}</strong> species</span>
      "))
    )
  )
}

icon_lookup <- tibble::tribble(
  ~iconic_group,        ~icon_path,
  "Brown algae",        "icons/Brown algae.png",
  "Crustaceans",        "icons/Crustaceans.png",
  "Ray-finned Fishes",  "icons/Fishes.png",
  "Green algae",        "icons/Green algae.png",
  "Molluscs",           "icons/Molluscs.png",
  "Red algae",          "icons/Red algae.png",
  "Segmented Worms",    "icons/Segmented worms.png",
  "Sponges",            "icons/Sponges.png",
  "Echinoderms",        "icons/Starfish.png",
  "Tunicates",          "icons/Tunicates.png",
  "Sea Anemones",           "icons/Anemones.png"
)
```

```{r q-announce}
div(
  class = "signature-intro-card",

  div(
    class = "signature-intro-icon",
    HTML("&#9733;")
  ),

  div(
    class = "signature-intro-content",

    div(
      class = "signature-intro-label",
      "New this quarter"
    ),

    h2(
      "Introducing Signature Species",
      class = "signature-intro-title"
    ),

    tags$p(
      class = "signature-intro-text",
      "Every Big Rock Pool Challenge hub now has its own signature species."
    ),

    tags$p(
      class = "signature-intro-text",
      paste(
        "By comparing thousands of wildlife records from across the UK,",
        "we've identified the species that is most characteristic of each location.",
        "As you explore the hub summaries, look out for these featured species",
        "and discover what makes every shoreline unique."
      )
    )
  )
)
```


```{r overall-summary}

tagList(

  render_summary_card(
    "Summary of all project data collected to date (30/06/2026)",
    observations = nrow(BRPC_dat),
    contributors = n_distinct(BRPC_dat$user.login),
    species = n_distinct(na.omit(BRPC_dat$scientific_name))
  )
)
```
# Big Rock Pool Challenge Biodiversity Report - Q2 2026

This report provides an overview of Big Rock Pool Challenge data collected during Quarter 2 of 2026 (01/04/2026-30/06/2026).

```{r Q2-summary}
tagList(

  render_summary_card(
    "Quarter 2 2026 Summary",
    observations = nrow(Q2_base),
    contributors = n_distinct(Q2_base$user.login),
    species = n_distinct(na.omit(Q2_base$scientific_name))
  )
)

```

### Iconic Taxa in Q2 2026

Species are grouped into higher-level “iconic taxa” categories based on their taxonomic ancestry. This grouping simplifies complex species-level data into broader biological categories (e.g. molluscs, crustaceans, algae), allowing clearer interpretation of community structure across sites and time periods.


```{r Q2-iconic-taxa}

Q2_iconic <- Q2_base %>%
  mutate(
    ancestry_vec = strsplit(
      ifelse(is.na(taxon.ancestry), "", taxon.ancestry),
      "/"
    ),

    iconic_group = purrr::map_chr(ancestry_vec, function(x) {

      x <- suppressWarnings(as.integer(x))

      hit <- iconic_lookup$Common[
        iconic_lookup$iNat_code %in% x
      ]

      if (length(hit) == 0) "Other" else hit[1]
    })
  )

iconic_summary <- Q2_iconic %>%
  count(iconic_group, sort = TRUE)

iconic_summary <- iconic_summary %>%
  left_join(icon_lookup, by = "iconic_group")


```

```{r add-rarity-scores}

# ---------------------------------------
# Join rarity scores
# ---------------------------------------

Q2_iconic <- Q2_iconic %>%
  left_join(
    rarity_scores,
    by = c("scientific_name" = "name")
  )
```
```{r iconic-Q2-plot}
ggplot(
  iconic_summary,
  aes(
    x = reorder(iconic_group, n),
    y = n
  )
) +
  geom_col(
    fill = "#9ecae1",
    width = 0.72
  ) +
  coord_flip() +
  scale_y_continuous(labels = comma) +

  geom_image(
    aes(image = icon_path),
    by = "width",
    size = 0.06,
    nudge_y = max(iconic_summary$n, na.rm = TRUE) * 0.03
  ) +

  expand_limits(
    y = max(iconic_summary$n, na.rm = TRUE) * 1.12
  ) +

  labs(
    title = "Iconic Taxa Recorded During Q2 2026",
    x = NULL,
    y = "Observations"
  ) +

  theme_minimal(base_size = 13) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(size = 11),
    plot.title = element_text(face = "bold")
  )
```

### Quarter 2 Records by Hub
Each hub represents a local coastal site where volunteers return to record marine life over time. These locations form the backbone of the Big Rock Pool Challenge network.

Because each hub has been running for different lengths of time, and is explored by different groups of people, the number of records varies, but every contribution adds valuable information about rock pool communities at that site.

Hubs are arranged in alphabetical order.

```{r Q2-map, echo=FALSE}

uk <- ne_countries(
  scale = "medium",
  country = "United Kingdom",
  returnclass = "sf"
) %>%
  st_transform(3857)


# Assign Q2 observations to nearest hub
Q2_sf$hub <- hub_sf$hub[
  sf::st_nearest_feature(Q2_sf, hub_sf)
]


# Remove geometry for summaries
Q2_map_data <- sf::st_drop_geometry(Q2_sf)


# Hub summaries for Q2 only
Q2_hub_map_summary <- Q2_map_data %>%
  group_by(hub) %>%
  summarise(
    records = n(),
    participants = n_distinct(user.login),
    species = n_distinct(scientific_name),
    .groups = "drop"
  )


# Add summaries to hub points
hub_sf_Q2 <- hub_sf %>%
  left_join(
    Q2_hub_map_summary,
    by = "hub"
  )


# Add hub IDs for labels
hub_sf_Q2 <- hub_sf_Q2 %>%
  mutate(
    hub_id = row_number()
  )


ggplot() +

  geom_sf(
    data = uk,
    fill = "grey95",
    colour = "grey50"
  ) +

  geom_sf(
    data = Q2_sf,
    colour = "grey70",
    alpha = 0.25,
    size = 0.3
  ) +

  geom_sf(
    data = hub_sf_Q2,
    aes(size = records),
    colour = "#1f78b4",
    fill = "#56B4E9",
    shape = 21,
    stroke = 1.2
  ) +

  geom_sf_text(
    data = hub_sf_Q2,
    aes(label = hub_id),
    colour = "black",
    fontface = "bold",
    size = 3
  ) +

  scale_size_area(
    max_size = 18,
    name = "Observations"
  ) +

  labs(
    title = "Q2 2026 Records by Hub",
    subtitle = "Records collected between 01/04/2026 and 30/06/2026"
  ) +

  theme_void() +

  theme(
    legend.position = "right",

    plot.title = element_text(
      face = "bold",
      size = 16
    ),

    plot.subtitle = element_text(
      size = 11,
      margin = margin(b = 10)
    )
  )
```


```{r Q2-hub-table, results='asis', echo=FALSE}

Q2_hub_table <- Q2_base %>%

  group_by(hub) %>%

  summarise(
    Observations = n(),
    Contributors = n_distinct(user.login),
    Species = n_distinct(na.omit(scientific_name)),
    .groups = "drop"
  ) %>%

  arrange(hub)


kable(
  Q2_hub_table,
  col.names = c(
    "Hub",
    "Observations",
    "Contributors",
    "Species"
  ),
  align = "lccc"
) %>%

  kable_styling(
    bootstrap_options = c(
      "striped",
      "hover",
      "condensed"
    ),
    full_width = FALSE
  )
```

### Top Scoring species in Q2

During BioBlitz Battle games, all species recorded earn a score. These scores reflect how often this species is recorded in the overall Big Rock Pool Challenge dataset. The higher scoring species are those that are recorded less often within the project and may represent more unusual or less commonly observed species.

In this report, scores are given between 1-100, according to our current scoring system 2.1

The highest scoring species recorded during this period are presented below.

```{r all-notable-species, results='asis', echo=FALSE}

library(dplyr)

q2_top <- Q2_base %>%
  left_join(
    rarity_scores,
    by = c("scientific_name" = "name")
  ) %>%
  filter(!is.na(Score)) %>%
  group_by(scientific_name) %>%
  summarise(
    Score = max(Score, na.rm = TRUE),

    photo = first(na.omit(taxon.default_photo.medium_url)),
    photo_attribution = first(na.omit(taxon.default_photo.attribution)),
    photo_license = first(na.omit(taxon.default_photo.license_code)),

    recorder = first(user.login),
    hub = first(hub),

    .groups = "drop"
  ) %>%

  mutate(
    license_ok = photo_license %in% allowed_licenses,

    photo = ifelse(
      license_ok & !is.na(photo) & photo != "",
      photo,
      default_avatar
    ),

    attribution = ifelse(
      license_ok,
      photo_attribution,
      ""
    )
  ) %>%

  arrange(desc(Score)) %>%
  slice_head(n = 3)
```

```{r standout-species-render, results='asis', echo=FALSE}
library(htmltools)

render_species_cards <- function(dat, title_text) {

  cards <- lapply(seq_len(nrow(dat)), function(i) {

    row <- dat[i, ]

    div(
      style = "
        flex:1;
        min-width:260px;
        max-width:320px;
        background:#ffffff;
        border-radius:18px;
        padding:20px;
        box-shadow:0 3px 12px rgba(0,0,0,0.08);
        border:1px solid #e6e6e6;
        text-align:center;
      ",

tags$a(
  href = row$photo,
  target = "_blank",
  tags$div(
    style = sprintf(
      "width:100%%;height:140px;border-radius:10px;background-image:url('%s');background-size:cover;background-position:center;",
      row$photo
    )
  )
),

      # FIX: must come from row, not global
      tags$div(
        style = "font-size:10px;color:#777;margin-top:6px;",
        HTML(row$attribution %||% "")
      ),

      div(
        style = "
          display:inline-block;
          padding:8px 14px;
          border-radius:999px;
          background:#f5f5f5;
          font-size:15px;
          margin-bottom:14px;
        ",
        strong("Score: "),
        round(row$Score, 1)
      ),

      tags$p(
        style = "margin-bottom:6px;",
        strong("Observed by"), tags$br(),
        row$recorder
      ),

      tags$p(
        style = "margin-top:0;",
        strong("Hub"), tags$br(),
        row$hub
      )
    )
  })

  tagList(
    tags$h2(
      style = "margin-top:40px;margin-bottom:24px;text-align:center;",
      title_text
    ),
    div(
      style = "
        display:flex;
        gap:24px;
        justify-content:center;
        flex-wrap:wrap;
        margin-bottom:50px;
      ",
      cards
    )
  )
}

render_species_cards(
  q2_top,
  "Highest scoring species across all hubs in Quarter 2"
)
```
::: {style="height:20px;"}
:::


```{r notable_species}

notable_species <- Q2_iconic %>%
  filter(!is.na(Score)) %>%
  group_by(scientific_name) %>%
  summarise(
    Score = max(Score, na.rm = TRUE),
    observations = n(),
    photo = first(na.omit(taxon.default_photo.medium_url)),
    recorder = first(user.login),
    hub = first(hub),
    .groups = "drop"
  ) %>%
  arrange(desc(Score)) %>%
  slice_head(n = 10)
```
### Top contributors in Q2 2026

The Big Rock Pool Challenge is powered by people. Every observation and identification is a contribution to a shared understanding of coastal biodiversity.

Some participants frequently attend events and submit records, while others help out by identifying species through verification on iNaturalist- all contributions are equally valuable in building the dataset.

The top recorders and identifiers are presented below- thank you for your efforts!

```{r Q2-hub-recorder-summary}

Q2_recorders <- Q2_base %>%
  filter(!is.na(user.login)) %>%
  group_by(user.login) %>%
  summarise(
    records = n(),
    icon = first(na.omit(user.icon_url)),
    .groups = "drop"
  ) %>%
  arrange(desc(records)) %>%
  slice_head(n = 3) %>%
  mutate(
    icon = ifelse(is.na(icon) | icon == "", default_avatar, icon),
    records = comma(records),

    card_html = paste0(
      "<div style='
        flex:1;
        min-width:120px;
        max-width:180px;
        background:#fafafa;
        border-radius:14px;
        padding:14px;
        text-align:center;
        border:1px solid #e6e6e6;
      '>",

      "<img src='", icon, "'
        style='
          width:64px;
          height:64px;
          border-radius:50%;
          margin-bottom:10px;
          border:2px solid #ececec;
          object-fit:cover;
        '>",

      "<div style='font-weight:600;font-size:14px;'>",
      user.login,
      "</div>",

      "<div style='font-size:13px;color:#666;margin-top:4px;'>",
      records,
      " records</div>",

      "</div>"
    )
  )

div(
  style = "
    background:#ffffff;
    border-radius:18px;
    padding:20px;
    box-shadow:0 3px 12px rgba(0,0,0,0.08);
    border:1px solid #ececec;
    margin:20px 0;
  ",

  h3("Q2 Top Recorders", style = "text-align:center;margin-top:0;"),

  div(
    style = "
      display:flex;
      gap:16px;
      justify-content:center;
      flex-wrap:wrap;
    ",
    HTML(paste(Q2_recorders$card_html, collapse = ""))
  )
)
```

```{r Q2-top-identifiers}

Q2_identifiers <- Q2_base %>%
  filter(!is.na(identifications)) %>%
  select(identifications) %>%
  mutate(
    identifier = map(identifications, ~ {
      if (is.null(.x) || nrow(.x) == 0 || !"user.login" %in% names(.x)) {
        character(0)
      } else {
        .x$user.login
      }
    }),
    icon = map(identifications, ~ {
      if (is.null(.x) || nrow(.x) == 0 || !"user.icon_url" %in% names(.x)) {
        character(0)
      } else {
        .x$user.icon_url
      }
    })
  ) %>%
  select(identifier, icon) %>%
  unnest(c(identifier, icon)) %>%
  filter(!is.na(identifier), identifier != "") %>%
  group_by(identifier) %>%
  summarise(
    identifications = n(),
    icon = first(na.omit(icon)),
    .groups = "drop"
  ) %>%
  arrange(desc(identifications)) %>%
  slice_head(n = 3) %>%
  mutate(
    icon = ifelse(is.na(icon) | icon == "", default_avatar, icon),
    identifications = comma(identifications),

    card_html = paste0(
      "<div style='
        flex:1;
        min-width:120px;
        max-width:180px;
        background:#fafafa;
        border-radius:14px;
        padding:14px;
        text-align:center;
        border:1px solid #e6e6e6;
      '>",

      "<img src='", icon, "'
        style='
          width:64px;
          height:64px;
          border-radius:50%;
          margin-bottom:10px;
          border:2px solid #ececec;
          object-fit:cover;
        '>",

      "<div style='font-weight:600;font-size:14px;'>",
      identifier,
      "</div>",

      "<div style='font-size:13px;color:#666;margin-top:4px;'>",
      identifications,
      " identifications</div>",

      "</div>"
    )
  )

div(
  style = "
    background:#ffffff;
    border-radius:18px;
    padding:20px;
    box-shadow:0 3px 12px rgba(0,0,0,0.08);
    border:1px solid #ececec;
    margin:20px 0;
  ",

  h3("Q2 Top Identifiers", style = "text-align:center;margin-top:0;"),

  div(
    style = "
      display:flex;
      gap:16px;
      justify-content:center;
      flex-wrap:wrap;
    ",
    HTML(paste(Q2_identifiers$card_html, collapse = ""))
  )
)
```

```{r load-hub-logos}
hub_logos <- tibble::tribble(
  ~hub, ~logo_url,
  "Castle Beach", "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68e52dddfca602b7e70fd34f_Castle%20Beach_background.png",
  "Lee-on-the-Solent", "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68f7a42bf2ced5a97ff1d7cc_Lee-on-the-Solent_background.png",
  "Beadnell Haven", "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68ee485ce2c8ba073e90f031_Beadnell%20Haven_background.png",
  "Ovingdean", "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68e7b2e3ee0622ac5288d0ea_Ovingdean%20Beach_background.png",
  "Shoeburyness", "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/69023480819461010c5b7777_Shoeburyness_background.png",
  "Menai Bridge", "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68f66091a36dd2e5a7f195a6_Menai%20Bridge_background.png",
  "Mount Batten", "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68ee4a28c6ce42f665792558_Mount%20Batten_background.png",
  "Kilve Beach", "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68e7af1e8dbb099b8d5242b3_Kilve%20Beach_background.png",
  "East Sands", "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68e7b07fedeb13e53466355c_East%20Sands%20Beach_background.png"
)
```

```{r load-hub-images}
hub_site_photos <- tibble::tribble(
  ~hub, ~site_photo,

  "Mount Batten",
  "https://cdn.prod.website-files.com/6761e748382f7a483dc6c262/687a3d7146a339fab83b27e5_Plymouth_lead_volunteers.jpg",

  "Castle Beach",
  "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68e525f5687223457755b1ef_4%20-%20Alexander%20Lydon.png",

  "Menai Bridge",
  "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68f65e11c5ec88584d09a9f8_menai%20bridge%20site.jpg",

  "Ovingdean",
  "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68e7b2c1af58aa97a264790e_IMG_3893%20-%20Sophie%20Crocker.jpeg",

  "East Sands",
  "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68f662a74fb848dae79f6a46_WhatsApp%20Image%202025-10-20%20at%2017.24.27_a9ec7b97.jpg",

  "Kilve Beach",
  "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68e7ae6ad6b9191fe5bbf189_20250905_142406%20-%20Athena%20Allen%20(1).jpg",

  "Beadnell Haven",
  "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68f663d2083ef94524503f95_team%20compressed.jpg",

  "Lee-on-the-Solent",
  "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68f7a41bfa2bbfcef8002094_IMG_20251018_142731.jpg",

  "Shoeburyness",
  "https://cdn.prod.website-files.com/6761e748382f7a483dc6c2ee/68f664397fe026047360f36a_shoeburyness%20team.jpg"
)
```

```{r hub-iconic-plots, echo=FALSE, warning=FALSE, message=FALSE}

dir.create("hub_plots", showWarnings = FALSE)

hub_iconic_small <- Q2_iconic %>%
  count(hub, iconic_group) %>%
  group_by(hub) %>%
  mutate(prop = n / sum(n)) %>%
  left_join(icon_lookup, by = "iconic_group")

for(h in unique(hub_iconic_small$hub)) {

  dat <- hub_iconic_small %>%
    filter(hub == h)

  p <- ggplot(
    dat,
    aes(
  x = reorder(iconic_group, prop),
  y = prop
    )
  ) +

   geom_col(
  fill = "#9ecae1",
  show.legend = FALSE
) +

    geom_image(
      aes(image = icon_path),
      by = "width",
      size = 0.05,
      nudge_y = max(dat$prop) * 0.03
    ) +

    expand_limits(
      y = max(dat$prop) * 1.12
    ) +

    coord_flip() +

    scale_y_continuous(labels = scales::percent) +

    labs(
      x = NULL,
      y = NULL
    ) +

    theme_minimal(base_size = 10) +

    theme(
      plot.margin = margin(2,2,2,2),
      panel.grid.minor = element_blank(),
      axis.text.y = element_text(size = 8),
      axis.text.x = element_text(size = 7)
    )

  ggsave(
    filename = paste0("hub_plots/", safe_id(h), ".jpg"),
    plot = p,
    width = 4,
    height = 2.8,
    dpi = 120
  )
}
```

### National BioBlitz 2026

The National BioBlitz 2026 was a UK-wide citizen science initiative coordinated by The Rock Pool Project, bringing people together to explore and record biodiversity on rocky shores. This annual event is a part of The Big Rock Pool Challenge. 

Held between 23rd and 31st May 2026, it invited individuals, families, and wildlife groups to survey intertidal habitats and contribute data to marine research and conservation.

As this event took place during the Q2 recording period, the data recorded are included in this report. Full details about the event, including the event report exploring data collected, can be found here: https://www.therockpoolproject.co.uk/national-bioblitz


During the National BioBlitz, **3,781** records were collected by **98** observers, of **544** species. 

Below is a map showing the location of all records collected during the 2026 National BioBlitz.

``` {r natbio}

NBioBlitz <- BRPC_dat %>%
  filter(
    observed_on >= as.Date("2026-05-23"),
    observed_on <= as.Date("2026-05-31")
  )
```

```{r natbioblitz-map}

library(sf)
library(rnaturalearth)

NBioBlitz_sf <- NBioBlitz %>%
  mutate(
    lon = purrr::map_dbl(geojson.coordinates, 1),
    lat = purrr::map_dbl(geojson.coordinates, 2)
  ) %>%
  filter(!is.na(lon), !is.na(lat)) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326)

uk_map <- ne_countries(
  scale = "medium",
  returnclass = "sf"
) %>%
  filter(admin %in% c("United Kingdom", "Ireland"))

ggplot() +
  geom_sf(
    data = uk_map,
    fill = "grey95",
    colour = "grey60"
  ) +
  geom_sf(
    data = NBioBlitz_sf,
    colour = "#0B6EF5",
    alpha = 0.7,
    size = 1
  ) +
  coord_sf(
    xlim = c(-11, 3),
    ylim = c(48.5, 61),
    expand = FALSE
  ) +
  labs(
    title = "National BioBlitz 2026 records"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(
      face = "bold",
      hjust = 0.5
    )
  )
```

### Hub Summaries

A breakdown of records collected at each Big Rock Pool Challenge hub is presented below, including key contributors, iconic taxa grouping, and top scoring species.


```{r Q2-hub-summary}

Q2_hub_summaries <- Q2_base %>%
  group_by(hub) %>%
  summarise(
    records = n(),
    contributors = n_distinct(user.login),
    species = n_distinct(na.omit(scientific_name)),
    .groups = "drop"
  )

```

```{r Q2-hub-top-groups}

Q2_hub_top_groups <- Q2_iconic %>%
  count(hub, iconic_group, sort = TRUE) %>%
  group_by(hub) %>%
  slice_head(n = 3) %>%
  summarise(
    top_groups = paste(iconic_group, collapse = ", "),
    .groups = "drop"
  )

```

```{r Q2-hub-top-species}
Q2_hub_top_species <- Q2_iconic %>%
  filter(!is.na(Score)) %>%

  group_by(hub, scientific_name) %>%

  arrange(desc(Score)) %>%

  slice(1) %>%

  ungroup() %>%

  transmute(
    hub,
    scientific_name,
    Score,

    photo = taxon.default_photo.medium_url,
    photo_attribution = taxon.default_photo.attribution,
    photo_license = taxon.default_photo.license_code,

    recorder = user.login
  ) %>%

  mutate(
    license_ok = photo_license %in% allowed_licenses,

    photo = ifelse(
      license_ok & !is.na(photo) & photo != "",
      photo,
      default_avatar
    ),

    attribution = ifelse(
      license_ok &
      !is.na(photo_attribution) &
      photo_attribution != "",
      photo_attribution,
      ""
    )
  ) %>%

  arrange(hub, desc(Score)) %>%

  group_by(hub) %>%

  slice_head(n = 3) %>%

  mutate(
    img_html = paste0(
      "<div style='
  flex:0 0 220px;
  width:220px;
  background:#ffffff;
        border-radius:14px;
        padding:14px;
        box-shadow:0 2px 8px rgba(0,0,0,0.08);
        border:1px solid #ececec;
        text-align:center;
      '>",

      "<img src='", photo, "' style='
          width:100%;
          height:140px;
          object-fit:cover;
          border-radius:10px;
          margin-bottom:6px;
      '>",

      "<div style='font-size:10px;color:#777;margin-bottom:8px;'>",
      attribution,
      "</div>",

      "<div style='font-style:italic;font-size:14px;margin-bottom:8px;'>",
      scientific_name,
      "</div>",

      "<div style='font-size:12px;color:#666;margin-bottom:8px;'>",
      "Rarity score: ", round(Score, 1),
      "</div>",

      "<div style='font-size:12px;color:#444;line-height:1.4;'>",
      "<strong>Observed by</strong><br>",
      recorder,
      "</div>",

      "</div>"
    )
  ) %>%

 summarise(
  top_species_html = paste0(
    "<div style='
display:flex;
justify-content:center;
align-items:flex-start;
gap:24px;
flex-wrap:wrap;
width:100%;
'>",
    paste(img_html, collapse = ""),
    "</div>"
  ),
  .groups = "drop"
)
```

```{r Q2-hub-top-recorders}

Q2_hub_top_recorders <- Q2_iconic %>%
  filter(!is.na(user.login)) %>%
  group_by(hub, user.login) %>%
  summarise(
    records = n(),
    icon = first(user.icon_url),
    .groups = "drop"
  ) %>%
  group_by(hub) %>%
  arrange(desc(records)) %>%
  slice_head(n = 3) %>%
  mutate(

    icon = ifelse(is.na(icon) | icon == "", default_avatar, icon),

    card_html = paste0(
      "<div style='
        flex:1;
        min-width:120px;
        max-width:180px;
        background:#fafafa;
        border-radius:14px;
        padding:14px;
        text-align:center;
      '>",

      "<img src='", icon, "'
        style='
          width:64px;
          height:64px;
          border-radius:50%;
          object-fit:cover;
          margin-bottom:10px;
          border:2px solid #e6e6e6;
        '>",

      "<div style='font-weight:600;font-size:14px;'>",
      user.login,
      "</div>",

      "<div style='font-size:13px;color:#666;margin-top:4px;'>",
      comma(records),
      " observations</div>",

      "</div>"
    )
  ) %>%
  summarise(
    top_recorders_html = paste(card_html, collapse = ""),
    .groups = "drop"
  )

```

```{r Q2-hub-top-identifiers}

Q2_hub_top_identifiers <- Q2_iconic %>%
  filter(!is.na(identifications)) %>%
  select(hub, identifications) %>%
  tidyr::unnest(identifications) %>%
  filter(!is.na(user.login)) %>%
  group_by(hub, user.login) %>%
  summarise(
    identifications = n(),
    icon = first(user.icon_url),
    .groups = "drop"
  ) %>%
  group_by(hub) %>%
  arrange(desc(identifications)) %>%
  slice_head(n = 3) %>%
  mutate(

    icon = ifelse(is.na(icon) | icon == "", default_avatar, icon),

    card_html = paste0(
      "<div style='
        flex:1;
        min-width:120px;
        max-width:180px;
        background:#fafafa;
        border-radius:14px;
        padding:14px;
        text-align:center;
      '>",

      "<img src='", icon, "'
        style='
          width:64px;
          height:64px;
          border-radius:50%;
          object-fit:cover;
          margin-bottom:10px;
          border:2px solid #e6e6e6;
        '>",

      "<div style='font-weight:600;font-size:14px;'>",
      user.login,
      "</div>",

      "<div style='font-size:13px;color:#666;margin-top:4px;'>",
      comma(identifications),
      " identifications</div>",

      "</div>"
    )
  ) %>%
  summarise(
    top_identifiers_html = paste(card_html, collapse = ""),
    .groups = "drop"
  )

```

```{r Q2-top-species-card}
Q2_top <- Q2_iconic %>%
  filter(!is.na(Score)) %>%
  group_by(scientific_name) %>%
  summarise(
    n = n(),

    Score = max(Score, na.rm = TRUE),

    photo = first(na.omit(taxon.default_photo.medium_url)),
    photo_attribution = first(na.omit(taxon.default_photo.attribution)),
    photo_license = first(na.omit(taxon.default_photo.license_code)),

    recorder = first(user.login),
    hub = first(hub),

    .groups = "drop"
  ) %>%

  mutate(
    license_ok = photo_license %in% allowed_licenses,

    photo = ifelse(
      license_ok & !is.na(photo) & photo != "",
      photo,
      default_avatar
    ),

    attribution = ifelse(
      license_ok,
      photo_attribution,
      ""
    )
  ) %>%

  arrange(desc(Score)) %>%
  slice_head(n = 3)

```


```{r Q2-hub-report}

Q2_hub_report <- Q2_hub_summaries %>%
  left_join(Q2_hub_top_recorders, by = "hub") %>%
  left_join(Q2_hub_top_identifiers, by = "hub") %>%
  left_join(Q2_hub_top_groups, by = "hub") %>%
  left_join(Q2_hub_top_species, by = "hub") %>%
  left_join(hub_logos, by = "hub") %>%
  left_join(hub_site_photos, by = "hub") %>%
  distinct(hub, .keep_all = TRUE)

```

```{r survey-hub-lookup}


survey_lookup <- BRPC_dat %>%

  mutate(
    survey_id = paste(
      user.login,
      observed_on,
      sep = "_"
    )
  ) %>%

  filter(
    !is.na(survey_id),
    !is.na(hub)
  ) %>%

  group_by(survey_id) %>%

  summarise(
    recorder = first(user.login),
    survey_date = first(observed_on),
    hub = first(hub),
    observations = n(),
    species = n_distinct(scientific_name),
    .groups = "drop"
  )
```

```{r key_species_matrix}

survey_species_matrix <- BRPC_dat %>%
  mutate(
    survey_id = paste(
      user.login,
      observed_on,
      sep = "_"
    )
  ) %>%
  filter(
    !is.na(scientific_name),
    !is.na(survey_id)
  ) %>%
  distinct(
    survey_id,
    scientific_name
  ) %>%
  mutate(
    presence = 1
  ) %>%
  pivot_wider(
    names_from = scientific_name,
    values_from = presence,
    values_fill = 0
  )

```

```{r matrix_lookup_hub}
survey_species_matrix_hub <- survey_species_matrix %>%
  left_join(
    survey_lookup,
    by = "survey_id"
  )

# Objects required for the separate signature-species analysis
signature_species_inputs <- list(
  survey_lookup = survey_lookup,
  survey_species_matrix = survey_species_matrix,

  # Only retain the taxonomic fields needed to create the lookup table
  taxon_data = BRPC_dat %>%
    dplyr::select(
      taxon.name,
      taxon.rank,
      taxon.id,
      taxon.preferred_common_name,
      taxon.default_photo.medium_url,
      taxon.default_photo.attribution,
      taxon.default_photo.license_code,
      photos,
      hub
    ) %>%
    dplyr::distinct()
)

# Create the output directory if necessary
dir.create(
  "analysis_data",
  showWarnings = FALSE,
  recursive = TRUE
)

saveRDS(
  signature_species_inputs,
  file = "analysis_data/signature_species_inputs.rds"
)

```

```{r load-sig-species-res}

signature_results_file <- "analysis_data/signature_species_results.rds"

if (!file.exists(signature_results_file)) {
  stop(
    "Signature species results not found.\n",
    "Run run_signature_species_analysis.R first."
  )
}

signature_species_outputs <- readRDS(signature_results_file)

signature_species <- signature_species_outputs$signature_species
signature_species_cards <- signature_species_outputs$signature_species_cards
signature_species_all_results <- signature_species_outputs$all_results
signature_species_hub_tables <- signature_species_outputs$hub_tables

signature_species_cards <- signature_species_cards %>%
  mutate(
    image_url = coalesce(image_url, default_avatar),
    image_attribution = coalesce(image_attribution, "")
  )

```



```{r Q2-hub-render}
Q2_hubs <- Q2_hub_report$hub
Q2_hubs <- Q2_hubs[!is.na(Q2_hubs)]

out <- lapply(Q2_hubs, function(h) {

  row <- Q2_hub_report %>%
    filter(hub == h) %>%
    slice(1)
  
  sig <- signature_species_cards %>%
  filter(hub == h) %>%
  slice_head(n = 1)

  sig_display_name <- if (
    nrow(sig) == 1 &&
    !is.na(sig$common_name) &&
    sig$common_name != ""
  ) {
    sig$common_name
  } else if (nrow(sig) == 1) {
    sig$species
  } else {
    ""
  }

  div(
  class = "panel"
    ,

    # HEADER
    div(
      style = "text-align:center;margin-bottom:24px;",

      img(
        src = row$logo_url,
        onerror = paste0("this.src='", default_avatar, "';"),
        style = "
          width:180px;
          height:180px;
          object-fit:contain;
          margin-bottom:14px;
        "
      ),

      h2(h, style = "margin:0;font-size:28px;"),

tags$div(
  style = "
    font-size:13px;
    color:#666;
    margin-top:6px;
    margin-bottom:4px;
  ",
  "Quarter 2: 01/04/2026–30/06/2026"
),

      div(
        style = "
          display:inline-block;
          margin-top:14px;
          padding:10px 18px;
          border-radius:999px;
          background:#f5f5f5;
          font-size:14px;
        ",
        strong(comma(row$records)), " observations · ",
        strong(comma(row$contributors)), " contributors · ",
        strong(comma(row$species)), " species"
      )
    ),

    # GRID
   div(
  class = "two-column",

      # SITE PHOTO
      div(
  class = "report-card",

        img(
          src = row$site_photo,
          onerror = paste0("this.src='", default_avatar, "';"),
          style = "
            width:100%;
            height:320px;
            object-fit:cover;
            border-radius:12px;
          "
        )
        ),

      # ICONIC TAXA
      div(
        style = "
          background:#fafafa;
          border-radius:18px;
          padding:18px;
          border:1px solid #ececec;
          text-align:center;
        ",

        h3("Iconic Taxa Composition", style = "margin-top:0;"),

        img(
          src = paste0("hub_plots/", safe_id(h), ".jpg"),
          onerror = paste0("this.src='", default_avatar, "';"),
          style = "
            width:100%;
            max-width:480px;
            border-radius:10px;
          "
        ),

        tags$p(
          style = "margin-top:16px;font-size:13px;color:#555;",
          strong("Top iconic groups: "),
          row$top_groups
        )
      ),

      # TOP RECORDERS
      div(
        style = "
          background:#fafafa;
          border-radius:18px;
          padding:18px;
          border:1px solid #ececec;
        ",
        h3("Top Recorders", style = "text-align:center;"),
        div(style="display:flex;gap:16px;justify-content:center;flex-wrap:wrap;",
            HTML(row$top_recorders_html))
      ),

      # TOP IDENTIFIERS
      div(
        style = "
          background:#fafafa;
          border-radius:18px;
          padding:18px;
          border:1px solid #ececec;
        ",
        h3("Top Identifiers", style = "text-align:center;"),
        div(style="display:flex;gap:16px;justify-content:center;flex-wrap:wrap;",
            HTML(row$top_identifiers_html))
      )
    ),

# SIGNATURE SPECIES
if (nrow(sig) == 1) {

  div(
    style = "
      position:relative;
      overflow:hidden;
      margin-top:30px;
      border-radius:22px;
      min-height:300px;
      box-shadow:0 8px 24px rgba(0,0,0,0.16);
      background:#123b52;
      color:#ffffff;
    ",

    # Background image
    img(
      src = sig$image_url,
      onerror = paste0("this.src='", default_avatar, "';"),
      style = "
        position:absolute;
        inset:0;
        width:100%;
        height:100%;
        object-fit:cover;
        object-position:center;
      "
    ),

    # Dark gradient overlay
    div(
      style = "
        position:absolute;
        inset:0;
        background:
          linear-gradient(
            90deg,
            rgba(5,25,38,0.96) 0%,
            rgba(5,25,38,0.86) 42%,
            rgba(5,25,38,0.30) 72%,
            rgba(5,25,38,0.12) 100%
          );
      "
    ),

    # Content
    div(
      style = "
        position:relative;
        z-index:2;
        padding:34px;
        max-width:560px;
      ",

      div(
        style = "
          display:inline-block;
          padding:7px 13px;
          border-radius:999px;
          background:rgba(255,255,255,0.16);
          border:1px solid rgba(255,255,255,0.25);
          font-size:12px;
          font-weight:700;
          letter-spacing:0.8px;
          text-transform:uppercase;
          margin-bottom:16px;
        ",
        paste0(
        toupper(h),
        " SIGNATURE SPECIES"
        )
      ),
      
      h3(
        sig_display_name,
        style = "
          margin:0 0 5px 0;
          font-size:30px;
          line-height:1.15;
          color:#ffffff;
        "
      ),

      tags$div(
        style = "
          font-size:18px;
          font-style:italic;
          margin-bottom:22px;
          color:rgba(255,255,255,0.88);
        ",
        sig$species
      ),

     tags$p(
      style = "
        margin:0;
        max-width:500px;
        font-size:14px;
        line-height:1.55;
        color:rgba(255,255,255,0.92);
      ",
      strong(h),
      " accounts for ",
      strong(comma(sig$observations)),
      " of the ",
      strong(comma(sig$overall_observations)),
      " Big Rock Pool Challenge records for this species."
    )
    ),

    # Attribution
    if (!is.na(sig$image_attribution) &&
        sig$image_attribution != "") {

      div(
        style = "
          position:absolute;
          z-index:3;
          right:12px;
          bottom:10px;
          max-width:45%;
          padding:5px 8px;
          border-radius:7px;
          background:rgba(0,0,0,0.58);
          color:rgba(255,255,255,0.85);
          font-size:9px;
          line-height:1.25;
          text-align:right;
        ",
        paste0(
          sig$image_attribution,
          if (
            !is.na(sig$image_licence) &&
            sig$image_licence != ""
          ) {
            paste0(" (", toupper(sig$image_licence), ")")
          } else {
            ""
          }
        )
      )
    }
  )

} else {

  NULL
},

    # SPECIES
    div(
      style = "margin-top:28px;",
      h3("Top Scoring Species", style = "text-align:center;margin-bottom:18px;"),
      HTML(row$top_species_html)
    )


  )
})

tagList(out)
```
